---
title: "Keep information from original dataset after a mapEdit in R"
author: |
  | Matthew Ross
  | Colorado State University
date: "10/17/2018"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r,setup}

knitr::opts_chunk$set(echo=T, warning=FALSE, message=FALSE)
```


I'm trying to snap points to a network (like roads or streams). Because many points need to be manually checked when snapping, it is easiest to just do this manually and use site metadata to determine where points should be moved to sit on the network. I'm hoping to do this all in R, but as far as I can tell the excellent package `mapedit` does not allow you to edit an exisiting feature and keep the original features' data (like an ID). This makes editing with `mapedit` less useful because I can see no way to join the snapped points back to their original information (other than a clunky spatial join, that would introduce as many problems as solutions).

Is this behavior expected? Am I doing something really simple that makes this not work? Any other R solutions that folks know of? 


# Simple example 

## Making up data
```{r}
library(sf)
library(mapview)
library(mapedit)
library(tidyverse)

#Make up some data near a road.
#I want to snap this data to the road

untethered <- tibble(lat=c(35,35.005),
               long=c(-80,-80.0001),
               id=c(1,2),
               road_name=c(1711,1712)) %>%
  st_as_sf(.,coords=c('long','lat'),crs=4326)

untethered
```

## Map raw data
```{r}
#Snap the raw data to the road names
map <- mapview(untethered)

map
```

## Snapping points
```{r,eval=F}
#Move points 
manually.snapped <- editMap(map,targetLayerId='untethered')
save(manually.snapped,file='snapped.RData')
```

## Check data
```{r}
load('snapped.RData')
mapview(manually.snapped$edited)

manually.snapped$edited
```

# Problem

How do I join the `manually.snapped` data back to the `untethered` data without a spatial join?